---
title: "The Effects of Marijuana Legalization on Crime and Other Drug Usage in Seattle"
author: "Benjamin Gaiser and Jeremy Russell"
date: "2 December 2016"
output: beamer_presentation
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Effects of Marijuana Legalization on Crime and Other Drug Usage in Seattle

Critics claim that **legalization** leads to an **increase in marijuana and other drug use**, that it "*increases crime, diminishes traffic safety, harms public health, and lowers teen educational achievement*." 

## Table of Contents

- Introduction\vspace{12pt}
- Research Question\vspace{12pt}
- Policy Relevance\vspace{12pt}
- Analysis\vspace{12pt}
- Conclusion\vspace{12pt}

## Introduction 
- Wanted to research crime and effects of policymaking on crime\vspace{12pt}

- Vox's article "Yes Weed Can" discussed upcoming legislation agenda on election day in November\vspace{12pt}

- Also mentioned two studies citing the inconclusive nature of current research given limited data\vspace{12pt}

- Data was indeed hard to find for a number of reasons, but decided on "Washington Initiative 502"\vspace{12pt}


## Research Question

- ***Does legalization of cannabis lead to an increase in crime?***\vspace{12pt}

- We look at different levels of crime (Alcohol, Burglary/Theft, Marijuana Usage, Other Drug Usage, Property Crime, and Violent Crime)\vspace{12pt}

- Due to data constraints we focused on Seattle instead of WA\vspace{12pt}

## Policy Relevance

- Cato Institute policy paper, "Dose of Reality: The Effect of State Marijuana Legalizations" published in September\vspace{12pt}

- Study covers Colorado, Washington, Oregon, Alaska and the effects of legalization on other drug and alcohol use, health, suicides, road safety, school expulsions\vspace{12pt}

- Conclude that data provide little support for strong arguments made by either opponents or supporters\vspace{12pt}

- RAND Corporation paper, "Options and Issues Regarding Marijuana Legalization" published last year\vspace{12pt}

- Authors argue (1) there are other options than usual binary between prohibition and "regulate like alcohol" and (2) it is too soon to judge the repurcusions of legalization and more data needed\vspace{12pt}

## Analysis - Descriptive Statistics (I)

------------------------------------------------------------------------
Variables                       Operationalisation
------------------------------  ----------------------------------------
Producer / Processor            Count of Producers and Processors in WA
Retailer                        Count of Retailers in WA 

Excise Tax                      Sum of Tax paid by Retailers
Sales                           Sum of Revenue Made by Retailers

Crime Categories                Aggregates of Crime.Event.Clearance.Description from Seattle 911-Incident Database

Established                     0 if no dispensary in district; 1 otherwise

Poverty Rate                    Share of People < $15.000 / year

Average Adults                  0 if less than mean share of adults in district; 1 otherwise

Average Whites                  0 if less than mean share of whites in district; 1 otherwise

GEOID                           U.S. Census Districts

------------------------------------------------------------------------

## Analysis - Descriptive Statistics (II)

```{r, include = FALSE}
source("./Codes/Third-Step - Plot Preparation/PlotsPreparation.R")
```

- Business is booming in WA and Sales and Tax Revenues are increasing

```{r, echo = FALSE}
cowplot::plot_grid(washington.barplot, washington.barplot.e, labels=c("",""))
```

## Analysis - Descriptive Statistics (III)

- Crime concentrates in central Seattle. Retailers seem not to be within the vicinty

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Map2 <- qmap("seattle", zoom = 12, source="google", maptype="roadmap")
Map2 + scale_fill_gradient2(breaks=pretty_breaks(n=5), guide=guide_legend(title="Population Density", keywidth = 2)) +
  geom_point(data=CrimeOct2016, aes(x=lon, y=lat, colour = factor(Crime)), alpha=0.75, size=1.1) +
  geom_point(data=Seattle.Crime.Analysis, aes(x=lon.Ret, y=lat.Ret), color="dark red", size=3) +
  labs(colour = "Crime Categories Oct 2016") + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

## Analysis - Descriptive Statistics (IV)

- Jump in February 2014 could mean that there is a reporting error that flaws our result (not the case)
- Next slide will show the development in log-scale

```{r pressure, echo=FALSE}
crime.graph <- ggplot(Crime.lines, aes(Month, CrimePerThousand, group=Crime, colour = factor(Crime))) 
crime.graph <- crime.graph + geom_line() +
  xlab("Months") + ylab("Crime frequency per 1000 inhabitants") + theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(colour = "Crime Categories") +
  ggtitle("Crime rates development in Seattle")

crime.graph.log <- ggplot(Crime.lines, aes(Month, log, group=Crime, colour = factor(Crime))) 
crime.graph.log <- crime.graph.log + geom_line() +
  xlab("Months") + ylab("Natural logarithm of Crime Rate per 1000 inhabitants") + theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(colour = "Crime Categories") +
  ggtitle("Crime rates development in Seattle")

crime.graph
```

## Analysis - Descriptive Statistics (V)

- Crimes in Property, Burglary/Theft, and Other Drugs seem to have changed most
```{r, echo = FALSE, message = FALSE, warning = FALSE}
crime.graph.log
```

## Analysis - Inferential Statistics (I)

- t-test shows that *treatment* differs with crime category

```{r, echo = FALSE, message = FALSE, warning = FALSE}
p <- cowplot::plot_grid(Box3, Box4, Box5, Box6, Box7, Box8, Box9, labels=c(""))
title <- ggdraw() + draw_label("Development of Crime per 1000 citizens", fontface="bold")
cowplot::plot_grid(title, p, ncol = 1, rel_heights = c(0.1, 1)) # rel_heights values control title margins
```

## Analysis - Inferential Statistics (II)


```{r, include = FALSE, message = FALSE, warning = FALSE}
ps.model <- glm(Established ~ log(CrimePerThousand), data = Seattle.Crime.Analysis,
                family=binomial,na.action=na.omit)
#summary(ps.model) # Pscores calculated by logit where outcome = binary treatment status
# covariates to be chosen that related to both treatment and potential outcomes

# Next calculate the pscores for each observations having a retailer
Seattle.Crime.Analysis$pscore <- predict(ps.model, newdata = Seattle.Crime.Analysis, type = "response")

# Figure problem from last plot
par("mar")
par(mar=c(1,1,1,1))

# mod_match <- matchit(Established ~ CrimePerThousand, method = "nearest", data = Seattle.Crime.Analysis)
Crime.pscore <- Seattle.Crime.Analysis[Seattle.Crime.Analysis$pscore >= 0.0009535593 & Seattle.Crime.Analysis$pscore <=0.4070721037,]
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
M1 <- lm(log(CrimePerThousand) ~ Established, data = Seattle.Crime.Analysis)
M2 <- lm(log(CrimePerThousand) ~ Established, data = Crime.pscore)
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
ps.model2 <- glm(Established ~ log(CrimePerThousand) + share_poverty + AgeCat + RaceCat, data = Seattle.Crime.Analysis,
                family=binomial,na.action=na.omit)
#summary(ps.model2) # Pscores calculated by logit where outcome = binary treatment status
# covariates to be chosen that related to both treatment and potential outcomes

# Next calculate the pscores for each observations having a retailer
Seattle.Crime.Analysis$pscore2 <- predict(ps.model2, newdata = Seattle.Crime.Analysis, type = "response")

# Figure problem from last plot
par("mar")
par(mar=c(1,1,1,1))

# mod_match <- matchit(Established ~ CrimePerThousand, method = "nearest", data = Seattle.Crime.Analysis)
Crime.pscore2 <- Seattle.Crime.Analysis[Seattle.Crime.Analysis$pscore2 >= 0.0003133269 & Seattle.Crime.Analysis$pscore <= 0.1961000,]
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
M3 <- lm(log(CrimePerThousand) ~ Established + share_poverty + AgeCat + RaceCat, data = Seattle.Crime.Analysis)
M4 <- lm(log(CrimePerThousand) ~ Established + share_poverty + AgeCat + RaceCat, data = Crime.pscore2)
```

```{r, echo=FALSE, fig.height=3.75, fig.width=5, message=FALSE, results="asis"}
stargazer::stargazer(M1, M2,
                     title="Analysis of Crime",
                     dep.var.labels = c("ln(Crime Rate Per Thousand)"),
                     covariate.labels = c("Established"),
                     digits=2, type="latex", header=FALSE, no.space = TRUE, single.row = TRUE, ci=TRUE, ci.level=0.95, omit.stat=c("LL","ser","f"))
```

## Analysis - Inferential Statistics (III)

```{r, echo=FALSE, fig.height=3.75, fig.width=4, message=FALSE, results="asis"}
stargazer::stargazer(M3, M4,
                     title="Analysis of Crime",
                     dep.var.labels = c("ln(Crime Rate Per Thousand)"),
                     covariate.labels = c("Established", "Poverty Rate", "> Average Adults", "> Average Whites"),
                     digits=2, type="latex", header=FALSE, no.space = TRUE, single.row = TRUE, ci=TRUE, ci.level=0.95, omit.stat=c("LL","ser","f"))
```

## Conclusion (I)

- Model 1 and 3 are non-matched regressions\vspace{12pt}
- Model 2 and 4 are matched regressions\vspace{12pt}
- There is a significant association between Retailers and crime development overall in Seattle. 
- Once we match similar districts, this becomes negligible though (CI + R²)


## Conclusion (II)

- Qualification I: self-selection bias\vspace{12pt}
- Qualification II: This analysis is for all crimes without using spatial or temporal dummies\vspace{12pt}
- Once we account for Month: Point estimates increase\vspace{12pt}
- Once we account for GEOID: Point estimates decrease\vspace{12pt}
- Qualification III: Not with propensity score matching (computational restrictions)

## Conclusion (III)

\begin{centering}
\LARGE{\textbf{Thank you for your attention.}} 
\end{centering}

## Annex (I)

- Common support problem with the univariate regression

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# There seems to be a common support problem!
labs <- paste("Retailer in GEOID established:", c("Yes", "No"))
Seattle.Crime.Analysis %>%
  mutate(Established = ifelse(Established == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pscore)) +
  geom_histogram(color = "white", breaks=seq(0,1,by=0.025)) +
  facet_wrap(~Established) +
  xlab("Probability of having a Retailer in GEOID (univariate)") +
  theme_bw()
```

## Annex (II)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# There seems to be a common support problem!
labs <- paste("Retailer in GEOID established:", c("Yes", "No"))
Seattle.Crime.Analysis %>%
  mutate(Established = ifelse(Established == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pscore2)) +
  geom_histogram(color = "white", breaks=seq(0,1,by=0.025)) +
  facet_wrap(~Established) +
  xlab("Probability of having a Retailer in GEOID (multivariate)") +
  theme_bw()
```

## Annex (III)

------------------------------------------------------------------------
Crime Category        M3     M4                      
------------------  -----  -----
Alcohol              0.7     NA
Burglary/Theft       1.5     NA
Marijuana           -0.5    -2.0 
Narcotics            0.5     NA
Other-Drugs          0.5    -0.6 
Property             1.4     NA
Violent              1.2     NA

For all p < 0.01
------------------------------------------------------------------------